\chapter*{Acknowledgements}
\label{chap:ack}

% A conclusão deste trabalho, bem como da grande maior parte da minha vida académica não seria possível sem a ajuda de ...
I would like to personally thank both my supervisors, Professor Mário M. Freire
and Professor Pedro R. M. Inácio, for the opportunity to develop this project
because it expanded my knowledge on various fields. I also want to acknowledge
their suuport and motivation throughout this endeavour.

Secondly I have to show my gratitude to all of my family, my parents, Fernando
Gomes and Eugénia Ramos, for educating me and supporting me, I would also like
to extend my gratitude to my grandparents, António Ramos and Maria Luísa
Veríssimo, and my uncles, Pedro Ramos and Sofia Ramos. Without my family I
would probably be a very different person.

I also want to recognize all the professors with whom I had the chance to learn
from, each and every one made me learn something useful that will influecne the
way I develop systems as well as the solutions I am able to design because of
what I have learned throughout my academic path.

Of course a honorable mention will go to Guido van Rossum, creator of the Python
programming language, not just because this project was developed with it but
also because he developed an entire sub-sulture which is based on open-source
and sharing knowledge without which I could achieve the system described in the
report.

Finally I want to give recognition to my friends who have supported and
encouraged me. Special mention to my best friend Filipa Costa for actually
understanding my dissapearance for along time.
