\chapter{Related Works and Technologies Employed}
% Os titulos dados aos capítulos são meros exemplos. Cada relatório deve adequar-se ao projeto desenvolvido.
\label{chap:related-material}

\section{Introduction}
\label{chap2:sec:intro}
% Cada capítulo \underline{intermédio} deve começar com uma breve introdução onde é explicado com um pouco mais de detalhe qual é o tema deste capítulo, e como é que se encontra organizado (i.e., o que é que cada secção seguinte discute).
This chapter starts off by describing similar solutions found to the system that
will be described throughout the report. Afterwards, in section 2.2, we will
move on to introduce the underlying techonologies that were leveraged in the
conception of this project.

\section{Similar Solutions}
\label{chap2:sec:similar-sol}
There are some solutions available which tackle some of the functionaliaties of
the system descrbied throughout this report, but to the best of the autors
knowledge none of them perform the exact all of the functionalities as a whole.
Now we will move on to describe these similar solutions which are:
\begin{enumerate}
	\item Nagios
	\item Domestic Routers
	% \item Wireshark
	% \item Snort
\end{enumerate}

Without further adue we will present all th e similar solutions enumerated above
and talk about the differences between each one and the system being described
in this report.

\subsection{Nagios}
\label{chap2:sec:similar-sol:sub:nagios}
One very similar project is the Nagios software, and it consists of a network
infrastructure monitoring system. Developed mostly for comercial and industrial
environments. There is a special version made for the Raspberry Pi, called
NagiosPi, that enables the user to check all of the connected devices and
monitor their traffic they generate. The NagiosPi system has the following
interfaces:
\begin{enumerate}
	\item Nagios;
	\item Nconf;
	\item NagVis;
	\item PHPMyAdmin;
	\item RaspControl;
\end{enumerate}

The NagiosPi system permits that the netwrok administrator logs all of the
information it might find pertinent as well as monitor all of the traffic
circulating inside the network. Using these interfaces the users can monitor
specific ports and hosts, it could visualize the packets and some statistics.

One of the biggest differences to the project being described in this report is
that it is more focused towards the domestic environments, because Nagios has a
complex interface using the web browser and not the mobile application offered
by this project. One more difference is that Nagios exclusively handles
monitoring, but the system being developed also has firewall capabilities, which
means it could block and also notification capabilities for flagged protocols
and packets.

\subsection{Domestic Routers}
\label{chap2:sec:similar-sol:sub:routers}
Another similar solution that could be referenced is the domestic router which
people have in their homes. This router actually comes with a firewall already
minimally configured preventing that any connection which isn't started by a
device inside the domestic network is unable to access or even probe the devices
which are inside the domestic network.

As said before this is a simple firewall but make a lot of difference because it
prevents malicious attackers from getting information about the network itself
and any of the devices connected to it, as well as prevents the exploitation of
these devices becasue they need that device to access a resource on the Internet
in order to become infected. This might sound useless but the reason wy people
get infected by a computer virus is because they access a website or resource on
the Internet which spreads malware, so it is clear that the user inside the
domestic network needs to access the infectious resource or download it in order
to get infected by it, in other words, it needs to bring the infection to
himself and not be always vulnerable.

The main difference between the firewall system of domestic routers and the
system described in this project is that the domestic router filters incoming
traffic, packets and connections coming from the Internet to the domestic
network. In the case of this system the filtering is done on the outgoins
packets, so they complement each other, preventing the user from getting malware
and also making it possible to prevent any communication between the virus and
any other resource on the Internet.
Another difference is that our system can be easily configurable using the
mobile application as well as notify the user of flagged communications by the
devices covered within the rule sets, which the domestic router can't do.
% \begin{lstlisting}[caption=Trecho de código usado no projeto.]
% #include <stdio.h>

% int main(){
%   int i = 0;
%   for(i = 0; i < 100; i++)
%     printf("%d\n",i);
% }
% \end{lstlisting}

\section{Technologies Employed}
\label{chap2:sec:tech}
In this section we will introduce the technologies that are being used
especially the ones that will host the system's components. This section won't
talk about the technologies that are used for the implementation itself, that is
a matter for the next chapter. This section will introduce the Raspberry Pi
board, which will host the server part of the system, and in section 2.2.3 there
is a brief description of the Android Operating System.

\subsection{Raspberry Pi}
\label{chap2:sec:tech:rasp}
The Raspberry Pi is a small computer that comprises all of the vital component
that make it a computer, with a CPU, memory and input and output capabilities.
This tiny computer was choosen due to its affordable price, less then 45€. For
that price the resources available are not extremly performant but for the
development of this gateway it is enough due to the low-level nature of the data
being accessed.

This micro-computer will have installed the Operating System and also the
project itself which comprises a web service, a database, a firewall and a
notification scrip for emails.

%talk about the specs of the Raspberry Pi 2?
Since one of the goals of this system is to develop a configurable gateway for
domestic networks, it is vital that the device that runs such gateway would be
inexpensive and the Raspberry Pi is a perfect solution because of it's good
price to power ratio.

\subsection{Android}
\label{chap2:sec:tech:android}
Android is an operating system focused on mobile platforms, it is based on the
Linux Kernel, which in turn was developed using the C programming language. This
operating system is published and maintainned by Google Inc. and the Open
Handset Alliance (OHA). The operating system is able to run on devices with the
different kinds of processors as the ARM, MIPS and x86.

The Android operating system is the most widely used OS for mobile devices
today, in part because of the low bar of entry in terms of price but also
because the Android OS is optimized to work with devices that don't have a lot
of computational resources. Because of the low price range that can be found
with the devices that run this operating system it is the ideal environment in
which to develop the companion and configuration application. So this
application will communicate and transmit the user's intentions to the system
itself which is hosted in the Raspberry Pi.

The main and official language of development of applications for this mobile
operating system is Java which is one of the most widely known programming
languages in the world, this enables the production of more applications and
associating it with the low prices on most devices it makes the Android
operating system widely adopted by consumers all over the world.

% make a point that Android has inexpensive devices like smartphones and tables.

\section{Conclusions}
\label{chap2:sec:concs}
This chapter talked about some of the similar products and also the main
differentiating characteristics between these solutions and our project, like
the Nagios monitoring system, or the firewall of domestic routers.
In this chapter there was also a description of the main technologies being
leveraged in this project, the Raspberry Pi inexpensive computer and also the
Android mobile operating system.
